---
title: "PCA"
author: "William"
date: '2022-06-08'
output: html_document
---

Diab dataset represent attributes with women who have cancer and those who don't based on the attributes given.


```{r}
#calling the libraries
library(factoextra)
library(ggplot2)
library(corrplot)
#
#loading the dataset
diab<-read.csv("C:\\Users\\HP 840 G3 i7\\Downloads\\diabetes.csv",header=T)

```

```{r}
#previewing the 1st 6 rows
head(diab)
```
```{r}
#getting attribute records/values
a <-diab$Glucose
head(a)
```
```{r}
#using attach function
attach(diab)
```

```{r}
#getting the attributes value
a <-Glucose
head(a)
```
```{r}
#statistics summary using summary()
summary(diab)
```

```{r}
#stat using describe()
library(psych)

```

```{r}
describe(diab)
```


```{r}
#checking the shape(dimension)
dim(diab)
```

```{r}
#running our principal component
# scale=T(True) means we use correlation matrix to calculate the PCA. if scale= F(False) then we use covariance matrix to calculate PCA.
# we use covariance(scale=F)when all the unit of our attributes are the same.
#we use correlation(scale=T) when the unit of our attributes are different.
diab.pca=prcomp(diab,scale=T)
```


```{r}
#checking the possible things we can get from running principal component(PC)
var=get_pca_var(diab.pca)
var

```
Observations:

coord >>> coordinate of the variables
cor >> correlation
cos2 >> how well is the attribute represented in each PC.
contrib >> how well each attribute/variable contribte to PC. 


```{r}
#getting the Eigenvalues
eig.val=get_eigenvalue(diab.pca)
eig.val
```

observation: 

5 PC accounts for 77.48% of the data


```{r}
#Plotting the PC
fviz_eig(diab.pca, addlabels = T, ylim = c(0, 30))
```
Observation: It shows the % of each PC
depending on how confidence we're we can stop at PC3 which explains 58% of the variance.


```{r}
#Using visualization to check the contribution. this is difficult to interpret.
fviz_pca_var(diab.pca, col.var = "blue")
```


```{r}
#Plotting the contribution of attributes in each PC
fviz_contrib(diab.pca,choice = "var", axes = 1)
fviz_contrib(diab.pca, choice = "var", axes = 2)
fviz_contrib(diab.pca, choice = "var", axes = 3)
fviz_contrib(diab.pca, choice = "var", axes = 4)
```
Observation: 

Dim1: Glucose, outcome, bmi, and insulin contribute the most in PC1
dim2: Age, pregnancy, skin thickness and insulin contribute the most.  

other Dims can be checked too...

These can be checked using corr plot as well.



```{r}
#checking attribute contribution using corr plot
corrplot(var$cos2, is.corr=FALSE)# u can also use is.corr=F
```
Observation:

Dim1 is a good presentation of outcome, BMI, and glucos
Dim2 is a good presentation of age, skin thickness, and pregrancies
Dim3 is a good presentation of blodd presure
Dim4 is a good presentation of DiabetesPedigreeFunction

This can still be got using the below code
```{r}
#Attribute contribution
corrplot(var$contrib, is.corr=F) 
```

```{r}
#checking the PC values
#this is a bit messy and abit confusing to interpret thus people prefer using visual representaion alredy done above.
diab.pca
```

Observations:
PC1: a decrease in glucose, a decrease in bmi, and a decrease in insulin leads to a decrease in the persons outcome.
PC2: an increase in age, pregnancy we have decrease in skin thickness and a decrease in insulin. this means the values are correlated in some way(inverse correlation).
Pc3: there's a relationship btn bllod pressure, outcome and glucose. i.e. glose goes up when blood pressure comes down.
